---
title: "Final Project"
author: "David, Saixiao, Alex"
date: "2024-11-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", out.width = "70%")
```

```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(tidyr)
```

1.

```{r}
# install.packages("tidytuesdayR")
# library(tidytuesdayR)
# tt_available()
```

2.
links: https://github.com/WSJ/measles-data
https://www.wsj.com/graphics/school-measles-rate-map/

```{r}
measles <- read.csv("/cloud/project/data/all-measles-rates.csv")
#summary(measles)
```

```{r}
measles$type <- as.factor(measles$type)
```
```{r}
measles<-measles %>%
 mutate(mmr_category= ifelse(mmr >=95,">=95%","<95%"))
```


## Part 1 Data Cleaning 

The minimum MMR rate is -1.00, which is unrealistic. Given that there are no missing values in the original "mmr" data, it is likely that missing values were mistakenly assigned the value of -1.00.

```{r}
measles_type <- measles%>%
filter(type != "Kindergarten" & type != "",mmr>=10) %>% 
  mutate(school_type = fct_collapse(type,
  Public = c("BOCES", "Public"), 
  Private = c("Nonpublic", "Private"), 
  Charter = "Charter")) %>%
mutate(school_type = fct_drop(school_type))
```


## Part 2: Exploratory Data Analysis

The data is left-skewed as most schools with mmr rate higher than 90%. 
```{r}
ggplot(measles_type, aes(x = mmr)) +
  geom_histogram(binwidth = 5, fill = "lightblue", 
                 col = "black", lwd = 0.4) +
  labs(title = "Plot 1 Distribution of MMR Vaccination Rate",
       x = "Vaccination Rate (%)",
       y = "Frequency") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

```



## Part 3: Data Analysis

### 3.1 One proportion hypothesis test

1. Write the hypotheses.

$H_0: p_0 = 0.8$ (The proportion of elementary schools in the U.S. with an MMR vaccination rate of at least 95% is 80%.)

$H_A: p_0 < 0.8$ (The proportion of elementary schools in the U.S. with an MMR vaccination rate of at least 95% is less than 80%.)

2. Check conditions.

1. Independence condition is satisfied since we assume all the elementary schools are independent.

2. Success-failure conditions is satisfied since:  

$np_0 =15130 * 0.80 = 12104 \geq 10$
$n(1 - p_0) = 15130 * 0.2 = 3026 \geq 10$


```{r}
addmargins(table(measles_type$mmr_category))
p0 <- 0.8
n1 <- 11769
n <- 15130
n*p0
n*(1-p0)
```
3. Test statistic

```{r}
# Hypothesis test for one-proportion
test1 <- prop.test(n1, n, p = 0.8, alternative = "less")
cat("Test statistic:", -sqrt(test1$statistic))
```

4. 
```{r}
p_val <- test1$p.value
cat("P-value =",p_val)
```


5. Decision and conclusion

Since p-value less than 0.05, we reject $H_0$, we have enough evidence that the true proportion of elementary schools in the US with MMR vaccination rate higher than 95% is less than 80%.


```{r}
cat("Confidence interval:", quantile(test1$conf.int, c(0.025, 0.975)), "\n")

```

### 3.2 Independence test

1. Hypotheses

$H_0:$ school type and MMR vaccination rate are independent.

$H_1$: school type and MMR vaccination rate are not independent.

2. Test conditions:

a. Independence: we are reasonable to assume all schools are independent with each other.

b. Expected counts: all greater than 5

```{r}
# Hypothesis test for independence
test2 <- chisq.test(measles_type$school_type, measles_type$mmr_category)
# Compute the expected count
cat("Expected counts:", test2$expected)
```

3. Test statistics

```{r}
cat("Test statistic:", test2$statistic)
```

4. P-value < 0.05

```{r}
cat("P-value:", test2$p.value)
```

5. Decision and conclusion

Decision: we reject $H_0$

Conclusion: we have enough evidence that school type and MMR vaccination rate are not independent with each other.



### 3.3 Exemption Analysis

```{r}
exemption_summary <- measles_type %>%
filter(xmed != -1 ,xper != -1 , xrel != -1)%>%
  group_by(school_type) %>%
  summarise(
    mean_med_exempt = mean(xmed, na.rm = TRUE),
    mean_pers_exempt = mean(xper, na.rm = TRUE),
    mean_rel_exempt = mean(xrel, na.rm = TRUE)
  )
exemption_summary
```

Limitations: Exemption data does not include charter schools.

Future Research: Investigating the reasons behind the lower vaccination rates in “charter schools” can provide actionable insights to policymakers.

